---
title: "Get started excluding data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(excluder)
```

The `{excluder}` package facilitates checking for, marking, and excluding rows of data frames for common online participant exclusion criteria. This package applies to online data with a focus on data collected from [Qualtrics](https://www.qualtrics.com) surveys, and default column names come from importing data with the  [`{qualtRics}`](https://docs.ropensci.org/qualtRics/) package. This may be most useful for [Mechanical Turk](https://www.mturk.com/) data to screen for duplicate entries from the same location/IP address or entries from locations outside of the United States.  However, the package can be used for data from other sources that need to be screened for IP address, location, duplicated locations, participant progress, survey completion time, and/or screen resolution.


## Usage

The package has three core actions:

1. `check` functions search for the exclusion criteria and output a message with the number of rows meeting the criteria and a data frame of the rows meeting the criteria. This is useful for viewing the potential exclusions.
1. `exclude` functions remove rows meeting the exclusion criteria. This is safest to do after checking the rows to ensure the exclusions are correct.
1. `mark` functions add a new column to the original data frame that labels the rows meeting the exclusion criteria. This is useful to label the potential exclusions for future processing without changing the original data frame.

The package provides seven types of exclusions based on Qualtrics metadata:

1. `duplicates` works with rows that have duplicate IP addresses and/or locations (latitude/longitude), using [`janitor::get_dupes()`](https://sfirke.github.io/janitor/reference/get_dupes.html).
1. `duration` works with rows whose survey completion time is too short and/or too long.
1. `ip` works with rows whose IP addresses are not found in the specified country (note: this exclusion type requires an internet connection to download the country's IP ranges), using package [`{iptools}`](https://github.com/hrbrmstr/iptools).
1. `location` works with rows whose latitude and longitude are not found in the United States.
1. `preview` works with rows that are survey previews.
1. `progress` works with rows in which the survey was not complete.
1. `resolution` works with rows whose screen resolution is not acceptable.

There are also helper functions:

1. [`collapse_exclusions()`](https://jeffreyrstevens.github.io/excluder/reference/collapse_exclusions.html)  collapses all of the columns marked by `mark` functions into a single column (each use of a `mark` function creates a new column).
1. [`deidentify()`](https://jeffreyrstevens.github.io/excluder/reference/deidentify.html) removes standard Qualtrics columns with identifiable information.
1. [`remove_label_rows()`](https://jeffreyrstevens.github.io/excluder/reference/remove_label_rows.html) removes the first two rows of labels and convert date and numeric columns in the metadata.

## Preparing your data
If you use the [`fetch_survey()`](https://docs.ropensci.org/qualtRics/reference/fetch_survey.html) from the `{qualtRics}` package to import your Qualtrics data, it will automatically remove the first two label rows from the data set. However, if you directly download your data from Qualtrics, it will include two rows in your data set that include label information. This has two side effects: (1) there are non-data rows that need to be removed from your data set, and (2) all of your columns will be imported as character data types.

The [`remove_label_rows()`](https://jeffreyrstevens.github.io/excluder/reference/remove_label_rows.html) function will remove these two label rows. Also, by default, it will coerce the Qualtrics metadata columns from character types to the correct formats (e.g., _StartDate_ is coerced to a date, _Progress_ is coerced to numeric). So if you download your data from Qualtrics, you will need to run this function on your data before proceeding.

```{r}
dplyr::glimpse(qualtrics_raw)
#
# Remove label rows and coerce metadata columns
df <- remove_label_rows(qualtrics_raw) %>% 
  dplyr::glimpse()
```


## Checking observations

Often the first step of excluding observations is to check which observations may meet a particular exclusion criterion. The `check_*()` suite of functions prints a message about the number of observations that meet the exclusion criterion and returns a data frame or tibble that includes only the observations that meet the criterion. 
```{r check1}
# Check for rows with incomplete data then rows with durations less than 100 seconds
df %>%
  check_progress()
```
```{r check2}
# Check for rows with with durations less than 100 seconds
df %>% 
  check_duration(min_duration = 100)
```

Because checks return only the rows meeting the criteria, they should not be connected via pipes unless you want to subset the second check criterion within the rows that meet the first criterion.

```{r check3}
# Check for rows with durations less than 100 seconds within rows that did not complete the survey
df %>%
  check_progress() %>%
  check_duration(min_duration = 100)
```


## Excluding observations
After checking which rows meet exclusion criteria, the `exclude_*()` suite of function will return a data frame that has removed observations that match the exclusion criteria. There is no need to pipe a check into an exclude, as the exclude functions already run the check. Therefore, excludes print a message of the number of observations meeting the criteria.

```{r exclude1}
# Exclude survey responses used to preview the survey
df %>% 
  exclude_preview() %>% 
  dplyr::glimpse()
```

Piping will apply subsequent excludes to the data frames with the previous excludes already applied. Therefore, it often makes sense to remove the preview surveys and incomplete surveys before checking other exclusion types to speed processing.

```{r exclude2}
# Exclude preview then incomplete progress rows then duplicate locations and IP addresses
df %>%
  exclude_preview() %>%
  exclude_progress() %>%
  exclude_duplicates()
```

## Marking observations
Rather than removing observations, you may want to simply mark them for future processing (or not). The `mark_*()` suite of functions creates a new column for each mark function used that marks which observations violate the exclusion criterion. Like the excludes, marks run the check functions, so they print a message about the number of observations meeting exclusion criteria. Mark functions return a data frame identical to the original with additional columns marking exclusions.

```{r mark1}
# Mark observations run as preview
df %>% 
  mark_preview() %>% 
  dplyr::glimpse()
```

Notice the new `exclusion_preview` column at the end of the data frame. It has marked the first two observations as `preview`.

Piping multiple mark functions will create multiple rows marking observations for exclusion.

```{r mark2}
# Mark preview and incomplete observations
df %>% 
  mark_preview() %>% 
  mark_progress() %>% 
  dplyr::glimpse()
```

To collapse all of the marked columns into a single column, use the `collapse_exclusions()` function. This will create a new `exclusions` columns that will collapse all exclusions in each observation into a single column. 

```{r mark3}
df %>% 
  mark_preview() %>% 
  mark_duration(min = 500) %>% 
  collapse_exclusions() %>% 
  dplyr::relocate(exclusions, .before = StartDate)
```

Multiple exclusions are separated by `,` but the separating character can be controlled by the `separator` argument. By default, the multiple exclusion columns are removed from the final data frame, but this can be turned off by setting the `remove` argument to `FALSE`. 

```{r mark4}
df %>% 
  mark_preview() %>% 
  mark_duration(min = 500) %>% 
  collapse_exclusions(separator = ";", remove = FALSE) %>% 
  dplyr::relocate(exclusions, .before = StartDate)
```


## Deidentifying data

By default, Qualtrics records participant IP address and location.^[To avoid recording this potentially identifiable information, go to _Survey options_ > _Security_ > _Anonymize responses_.] You can also record properties of the participants' computers such as operating system, web browser type and version, and screen resolution.^[You can turn these on by adding a new question (usually to the beginning of your survey) and choosing _Meta info_ as the question type.] While these pieces of information can be useful for excluding observations, they carry potentially identifiable information. Therefore, you may want to remove them from data frame before saving or processing it. The `deidentify()` function removes potentially identifiable data columns collected by Qualtrics. By default, the function uses a strict rule to remove IP address, location, and computer information (browser type and version, operating system, and screen resolution).

```{r deidentify1}
# Exclude preview then incomplete progress rows
df %>%
  exclude_preview() %>%
  exclude_progress() %>%
  exclude_duplicates() %>%
  deidentify() %>%
  dplyr::glimpse()
```

If the computer information is not considered sensitive, it can be kept by setting the `strict` argument to `FALSE`, thereby only removing IP address and location.

```{r deidentify2}
# Exclude preview then incomplete progress rows
df %>%
  exclude_preview() %>%
  exclude_progress() %>%
  exclude_duplicates() %>%
  deidentify(strict = FALSE) %>%
  dplyr::glimpse()
```

